---
title: "Seurat_Code_LD_2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r}
#install.packages("BiocManager")

#BiocManager::install("Seurat")
#BiocManager::install("ggplot2")
#BiocManager::install("sctransform")

library(Seurat)
library(ggplot2)
library(sctransform)
library(dplyr)
library(ape)
library(variancePartition)
library(cowplot)
library(Matrix)

#find out where you are
getwd()
dir= "/Users/whippoorwill/Desktop/Sequencing/Nick/2_2020"
 #set this to whatever you like, above the level of your data folder
#setwd(dir)
```

Select only P28 samples: 

```{r}
load("/Users/whippoorwill/Desktop/Sequencing/Nick/2_2020/Graphs_Analysis_August2020 mpeg only/Data/mpegonly_all.RData")
sobject
sobject = subset(sobject,subset = age == "Juvenile")
sobject
```

```{r}
sobject <- NormalizeData(sobject,normalization.method = "LogNormalize", scale.factor = 10000)
#select some housekeeping genes
genes = c("mpeg1.1",'hexb','ptprc') 
VlnPlot(sobject,genes,group.by = "sample_description")
GetAssayData(sobject)[1:10, 1:15]
table(sobject$sample_description)
```


Filtering out genes in Seurat - use AFTER normalization. Works on data object (i.e. normalized, filtered data). 


```{r}
sobject<-FindVariableFeatures(sobject, selection.method = "vst", nfeatures = 6000)
top10 <- head(VariableFeatures(sobject),10)
plot1 <- VariableFeaturePlot(sobject)
plot2 <- LabelPoints(plot=plot1, points = top10, repel = T, xnudge = 0, ynudge = 0)
plot2
```

Scale the data
```{r}
all.genes<-rownames(sobject)
var.genes = VariableFeatures(sobject)
add.genes = c('fcrls', 'olfml3', 'lgmn', 'cx3cr1', 'hexb','apoeb','mpeg1.1')
add.genes = add.genes[!add.genes %in% var.genes]
any(add.genes %in% var.genes)
scalegenes = c(var.genes,add.genes)

sobject<-ScaleData(sobject,features = scalegenes, vars.to.regress = c("nFeature_RNA"))
```

Dimensionality reduction. Note that UMAP can be run directly on most variable features as well, but this will take forever unless you severely restrict #genes. 
```{r}
sobject <- RunPCA(sobject,features = scalegenes,npcs = 50, verbose = FALSE)
set.seed(1)
#figure out how many PCs to use - can default to 30
ElbowPlot(sobject,ndims = 50, reduction = "pca")
print(sobject[["pca"]], dims = 1:20, nfeatures = 5)
#look for heat shock proteins or Malat1
FeaturePlot(sobject,"percent.mito",reduction = "pca",dims = 1:2 )
```


```{r}
sobject<-RunUMAP(sobject,reduction = "pca",dims = 1:30, verbose = F)
sobject<-FindNeighbors(sobject,dims=1:30,verbose=F)
set.seed(1)
res = seq(0.5,3,by = 0.5) #set this to anything
sobject<-FindClusters(sobject,verbose=F,resolution = 0.3) 
save(sobject,file = "~/Desktop/mpeg_juvenile.RData")
```


Function to print multiple graphs: 

```{r}
PrintSeuratGraph = function(namecard = "a",sobject,graphtype = "feature",feature = NULL,group = NULL,split=NULL,cellnames=NULL,sort = F){
  if (!is.null(cellnames)){
    Idents(sobject) = cellnames[1]
    cells = colnames(sobject)[Idents(sobject) %in% cellnames[2:length(cellnames)]]} 
  else {cells = cellnames}
  if (graphtype == "feature"){
    graph = FeaturePlot(sobject,features = feature,split.by = split, cells = cells,cols = c("lightgrey","lightblue","yellow","orange","darkorange","red"))
  }
  if (graphtype == "violin"){
    graph = VlnPlot(sobject,features = feature, pt.size = 0.1, idents = cellnames[2:length(cellnames)],group.by = group, split.by = split)
  }
  if (graphtype == "dim"){
    graph = DimPlot(sobject,cells = cells, group.by = group, split.by = split)
    
  }
  name = paste0(feature,"_",graphtype,namecard,".eps")
  graph
  setEPS()
  postscript(file.path("~/Desktop",name))
  print(graph)
  dev.off()
}

```

```{r}
DimPlot(sobject,label=T,reduction = "umap")
genes = c("marco","mpeg1.1","apoeb","p2ry12","ptprc","hexb","irf8","spi1b","spi1a","csf1ra","csf1rb","slc7a7","c1qa","c1qb","c1qc")

features = c("percent.mito","percent.ribo","nCount_RNA","nFeature_RNA")
genesHB = c("ccl19a.1","wipf1a","b2m","plp1b","krt18b","cxcr4b","cd74b","cd74a","gpatch8","mhc2dab","ms4a17a.6","cdc42l","dhrs9","bzw1b","lygl1","hmga1a","top1l","plekho2")

genesOT = c("bzw2","slc40a1","g0s2","ctsla","eepd1","slc43a2b","apoeb","pitpnaa","rnaset2","mt2","ctsz","cndp2","dpp7","abca1a","pdgfd","ctsd","pepd","mafbb","ppdpfb","slc7a7","glud1b","porb","galk1","bri3","glula","abhd12","ubb","rgl1","mpp1","hmgb3a","abca1b","ctsba","ctsc","vat1")

length(genesOT)



genes = genes[genes %in% scalegenes]
FeaturePlot(sobject,genes[1:5],reduction = "umap")
FeaturePlot(sobject,genes[6:9], reduction = "umap")
FeaturePlot(sobject,features = "percent.mito")
FeaturePlot(sobject,features = "nCount_RNA")
FeaturePlot(sobject,features = "nFeature_RNA")
FeaturePlot(sobject,features = "percent.ribo")
DimPlot(sobject,group.by = "individual")

genes = c("ctsba", "psap", "ctsla", "lygl1","c1qa","cebpb","g0s2","irf2")

for(feature in c(genes,features)){
  PrintSeuratGraph(namecard = "Juvenile_mpeg",sobject=sobject,graphtype = "feature",feature = feature)
}

for(feature in c(features)){
  PrintSeuratGraph(namecard = "Juvenile_mpeg_split",sobject=sobject,graphtype = "feature",feature = feature,split = "individual")
}

groups = c("sample_description","seurat_clusters")

for(group in groups){
  PrintSeuratGraph(namecard = "Juvenile_mpeg_individual",sobject=sobject,graphtype = "dim",group = group, feature = group,split = "individual")
}

genes = c("apoeb", "p2ry12", "hexb", "csf1rb", "irf8", "slc7a7","mpeg1.1","ptprc")
genes = c("siglec15l", "grn1", "flt3", "spock3", "actn1","ccl19a.1")

for (gene in genes){
  PrintSeuratGraph(namecard = "Juvenile_mpeg_vln",sobject=sobject,graphtype = "violin",group = "seurat_clusters", feature = gene)
}

DimPlot(sobject,group.by = "seurat_clusters",label = T)
```



```{r}
DimPlot(sobject,reduction = "umap",group.by = "sample_description")
DimPlot(sobject,reduction = "umap",group.by = "sample_description",split.by = "age")
setEPS()
postscript("~/Desktop/juvenile_clusters_mpeg.eps")
DimPlot(sobject,reduction = "umap",group.by = "seurat_clusters",label = T)
dev.off()

setEPS()
postscript("~/Desktop/juvenile_individual_mpeg.eps")
DimPlot(sobject,reduction = "umap",group.by = "individual",label = T)
dev.off()
```
#The clusters are stored in "seurat_clusters", with multiple clustering algorithms stored as metadata columns. 
```{r}
x = table(sobject$individual,sobject$seurat_clusters)
ratio = x[1,]/x[2,]*100
b = round(rbind(x,ratio),0)
b
mean(b[3,])
barplot(x,legend.text = T)
```

```{r}
Idents(sobject) = "seurat_clusters"
sobject= BuildClusterTree(sobject,dims = 1:30)
tree = sobject@tools$BuildClusterTree
setEPS()
postscript("~/Desktop/tree_juvenile_mpeg.eps")
plot.phylo(tree, use.edge.length = T, direction = "rightwards")
dev.off()
```

Heatmap for sobject
```{r}
Idents(sobject) = "seurat_clusters"
markers_all <- FindAllMarkers(
    object = sobject,
    features = rownames(GetAssayData(sobject,slot = "scale.data")),
    test.use = "MAST",
    only.pos = FALSE, 
    min.pct = 0.25, #gene must be present in 25% of the cells in the cluster
    logfc.threshold = 0.20
)

dim(markers_all)
head(markers_all)
write.csv(markers_all,file = "~/Desktop/juvenile_mpeg_markers.csv")

markers_all = markers_all[markers_all$avg_logFC>0,]
pcut = 10^-50
markers_all = markers_all[markers_all$p_val_adj<pcut,]
markers_all_single <- markers_all[markers_all$gene %in% names(table(markers_all$gene))[table(markers_all$gene) == 1],] #markers that define a single cluster



topgenes <- markers_all_single %>% group_by(cluster) %>% top_n(5, avg_logFC)

setEPS()
postscript("~/Desktop/heatmap_juvenilempeg_pval50.eps")
DoHeatmap(
    object = sobject, 
    features = topgenes$gene,
    group.by = 'seurat_clusters',
    size = 5,
    label = T,
    draw.lines = T
)

dev.off()
```
Find a gene that defines microglia vs all cells
